Make any stupid plot you want

### 2021-05-05 draft/sections/math.tex
q & qhat

\subsection{Artist}
\label{sec:artist}
The topological artist \vartist\ is how we model the building block component that transforms data into a graphic. The artist \vartist\ is a map from the sheaf on a data bundle \dtotal\ which is $\mathcal{O}(\dtotal)$ to the sheaf on the graphic bundle \gtotal, $\mathcal{O}(\gtotal)$. 
\begin{equation}
    \vartist: \mathcal{O}(\dtotal) \rightarrow \mathcal{O}(\gtotal)
    \label{eq:math:artist:artist}
\end{equation}
The artist preserves \textit{continuity} through the \vindex\ map discussed in \autoref{sec:math:graphic:base}. We conjecture that it is an \textit{equivariant} map because it carries a homomorphism of monoid actions \cite{cegarraCohomologyMonoidsOperators2019}
\begin{equation}
    \varphi: \monoid \rightarrow \monoid^{\prime}
\end{equation}

Given \monoid\ on data $\mathscr{\dtotal}$ and $\monoid^{\prime}$ on graphic $\mathscr{\gtotal}$, we propose that artists $\mathscr{\vartist}$ are equivariant maps 
\begin{equation}
\vartist(m\cdot \delement) = \varphi(m)\cdot \vartist(\delement) 
\end{equation}
such that applying a monoid action $m \in \monoid$ to the data input $\delement \in \mathscr{\dtotal}$ of the artist $\mathscr{\vartist}$ is equivalent to applying a monoid action $\varphi(\monoid) \in \monoid^{\prime}$ to the graphic $\vartist(\delement) \in \mathscr{\gtotal}$ output of the artist.

The monoid equivariant map has two stages: the encoders $\vchannel:\dtotal^{\prime} \rightarrow \vtotal$ convert the data components to visual components, and the assembly function $\vmark: \vtotalpull \rightarrow \gtotal$ composites the fiber components of \vtotalpull\ into a graphic in \gtotal.
\begin{equation}
    \label{eq:math:artist:diagram}
    \begin{tikzcd}
        \dtotal \arrow[r, "\vchannel"] \arrow[rd, "\pi"'] & \vtotal \arrow[d, "\pi"] & \vindex^*\vtotal \arrow[r, "\vmark"] \arrow[d, "\vindex^*\pi"'] \arrow[l, "\vindex^*"'] & \gtotal \arrow[ld, "\pi"] \\
                                              & \dbase                  & \gbase \arrow[l, "\vindex"']                                              &                    
        \end{tikzcd}
\end{equation}
\vtotalpull\ is the visual bundle \vtotal\ pulled back over \gbase\ via the equivariant continuity map $\vindex:\gbase \rightarrow \dbase$ introduced in \autoref{sec:math:graphic:base}.
The functional decomposition of the visualization artist in \autoref{eq:math:artist:diagram}facilitates building reusable components at each stage of the transformation because the equivariance constraints are defined on \vchannel, \vmark, and \vindex. We name this map the artist as that is the analogous part of the  Matplotlib\cite{hunterArchitectureOpenSource} architecture that builds visual elements.

\subsubsection {Visual Fiber Bundle \vtotal}
\label{sec:math:visual}
We introduce a visual bundle \vtotal\ to store the mappings of the data components into components of the graphic. The visual bundle $(\vtotal,\,\dbase,\,\pi ,\,\vfiber)$ is the space of possible parameters of a visualization type, such as a scatter or line plot. As with the data and graphic bundles, the visual bundle is defined by the projection map $\pi$
\begin{equation}
    \begin{tikzcd}[ampersand replacement=\&]
        \vfiber \arrow[r, hook] \& \vtotal \arrow[d, "\pi"'] \\
                          \& \dbase \arrow[u, "\vsection"', bend right]
    \end{tikzcd}
\end{equation}
where \vsection\ is the visual variable encoding, as described by Bertin \cite{bertinSemiologyGraphicsDiagrams2011a}, of the data section \dsection. The visual fiber \vfiber\ is defined in terms of the input parameters of the visualization library's plotting functions; by making these parameters explicit components of the fiber, we can build consistent definitions and expectations of how these parameters behave.
\begin{table}[H]
    \centering
    \renewcommand{\arraystretch}{2}
    \begin{tabulary}{\textwidth}{|l|L|l|}\hline
     $\bm{\vchannel_{i}}$                      & $\bm{\vsection_{i}}$                                                            & $\bm{codomain(\vchannel_{i}) \subset \vfiber_{i}}$  \\ \hline                                              
    position                    & x, y, z, theta, r                                                          & $\mathbb{R}$   \\ \hline
    size                        & linewidth, markersize                                            & $\mathbb{R}^{+}$   \\ \hline
    shape                       & markerstyle                                                      & $\{f_{0}, \ldots, f_{n}\}$ \\ \hline
    color                       & color, facecolor, markerfacecolor, edgecolor  & $\mathbb{R}^{4}$ \\ \hline
    \multirow{2}{*}{texture}    & hatch                                                            & $\mathbb{N}^{10}$\\\cline{2-3}
                                & linestyle                                                        & $(\mathbb{R}, \mathbb{R^+}^{n, n\%2=0})$ \\ \hline              
    \end{tabulary}
    \caption{Some possible components of the fiber \vfiber\ for a visualization function implemented in Matplotlib}
    \label{tab:math:artist:mpl:fiber}
\end{table}
 A section \vsection\ is a tuple of visual values that specifies the visual characteristics of a part of the graphic. For example, given a fiber of \(\{x, y, color\}\) one possible section could be  \(\{.5, .5, (255, 20,147)\}\). The \(codomain(\vchannel_i)\) determines which monoids can act on $\vfiber_i$. These fiber components are implicit in the library, as seen in \autoref{tab:math:artist:mpl:fiber}, and by making them explicit as components of the fiber we can build consistent definitions and expectations of how these parameters behave.

\subsubsection{Visual Encoders \vchannel}
\label{sec:math:artist:nu}
We define the visual transformers \vchannel\ 
\begin{equation}
  \label{eq:math:artist:nu}
  \{\vchannel_{0}, \ldots, \vchannel_{n}\}: \{\dsection_{0}, \ldots, \dsection_{n}\} \mapsto \{\vsection_{0}, \ldots, \vsection_{n}\}
\end{equation}
as the set of equivariant maps $\vchannel_i: \dsection_i \mapsto \vsection_i$. Given $\monoid_i$ is the monoid action on $\dtotal_i$ and that there is a monoid ${\monoid_i}^{\prime}$ on $\vtotal_i$, then there is a monoid homomorphism from $\varphi:\monoid_i \rightarrow {\monoid_i}^{\prime}$ that \vchannel\ must preserve. As mentioned in \autoref{sec:math:data:monoid}, monoid actions define the structure on the fiber components and are therefore the basis for equivariance. A validly constructed \vchannel\ is one where the diagram of the monoid transform $m$ commutes
\begin{equation}
  \label{eq:math:artist:nu_commute}
\begin{tikzcd}
  \dtotal_i \arrow[r] \arrow[r, "\vchannel_i"] \arrow[d, "m_{\delement}"'] & \vtotal_i \arrow[d, "m_{\velement}"] \\
  \dtotal_i \arrow[r, "\vchannel_i"]                           & \vtotal_i               
\end{tikzcd}
\end{equation}
such that applying equivariant monoid actions to $\dtotal_{i}$ and $\vtotal_{i}$ preserves the map $\vchannel_{i}: \dtotal_{i}\rightarrow \vtotal_{i}$. In general, the data fiber $\dfiber_{i}$ cannot be assumed to be of the same type as the visual fiber $\vfiber_{i}$ and the actions of \monoid\ on $\dfiber_{i}$ cannot be assumed to be the same as the actions of $\monoid^{\prime}$ on \vfiber; therefore an equivariant $\vchannel_i$ must satisfy the constraint  
\begin{equation}
\vchannel_i(m_{\delement}(\dtotal_i)) = \varphi(m_{\delement})(\vchannel_i(\dtotal_i))
\label{eq:math:artist:nu:equivariance}
\end{equation} 
such that $\varphi$ maps a monoid action on data to a monoid action on visual elements. However, we can construct a monoid action of \monoid\ on $\vfiber_{i}$ that is compatible with a monoid action of \monoid\ on $\dfiber_{i}$. We can compose the monoid actions on the visual fiber $\monoid^{\prime} \times \vfiber_{i} \rightarrow \vfiber_{i}$ with the homomorphism $\varphi$ that takes \monoid\ to $\monoid^{\prime}$. This allows us to define a monoid action on \vfiber\ of \monoid\ that is $(m, \velement) \rightarrow \varphi(m)\bullet\velement$. Therefore, without a loss of generality, we can assume that an action of \monoid\ acts on $\dfiber_{i}$ and on $\vfiber_{i}$ compatibly such that $\varphi$ is the identity function.

\begin{figure}[htb]
  \centering
  \begin{subfigure}{.49\textwidth}
    \includegraphics[width=1\textwidth]{figures/math/partial_invalid.png}
    \caption{}
    \label{fig:math:artist:nu:invalid}
  \end{subfigure}
  \begin{subfigure}{.49\textwidth}
    \includegraphics[width=1\textwidth]{figures/math/partial_fixed.png}
    \caption{}
    \label{fig:math:artist:nu:valid}
  \end{subfigure}
  \caption{The map from data component to visual component in \autoref{fig:math:artist:nu:invalid} is not homomorphic, and therefore invalid, because $rain \geq storm$ is mapped to elements with the reverse ordering $\nu(storm) \geq \nu(storm)$. In contrast, the mapping in \autoref{fig:math:artist:nu:valid} is valud since $\nu(storm)=\nu(rain)$ satisfies the condition $\nu(storm) \geq \nu(storm)$}
  \label{fig:math:artist:nu}
\end{figure}

The translation from weather state data to visual representation as umbrella emoji in \autoref{fig:math:artist:nu:invalid} is an invalid visual encoding map \vchannel\ because it is not homomorphic. This is because the monotonic condition \(rain \geq storm \implies \vchannel(rain) \geq  \vchannel(storm)\) is not met since \(\vchannel(rain) \leq\vchannel(storm)\). To satisfy the monotonic condition for $rain \geq storm$, either red arrow in \autoref{fig:math:artist:nu:invalid} would have to go in a different direction. On the other hand, the mapping from weather state to umbrellla in \autoref{fig:math:artist:nu:valid} is a homomorphism since $\vchannel(rain) = \vchannel(storm)$ satisfies the monotonic condition of $rain \geq storm$. \autoref{fig:math:artist:nu} is an example of how the model supports partially ordered data components, which was a motivation for defining equivariance as monoid homomorphisms.

\begin{table}[H]
\centering
  \renewcommand{\arraystretch}{2}
  \begin{tabulary}{\columnwidth}{|llL|}\hline
      \textbf{scale} & \textbf{group} & \textbf{constraint}\\ \hline
      nominal & permutation &  $\text{if } \delement_1 \neq \delement_2 \text{ then } \vchannel (\delement_1) \neq\vchannel(\delement_2)$\\
      ordinal &  monotonic & $\text{if } \delement_1 \leq \delement_2 \text{ then } \vchannel (\delement_1) \leq \vchannel(\delement_2)$\\
      interval &  translation &  $\vchannel (x + c) = \vchannel(x) + c$ \\
      ratio &  scaling &  $\vchannel(xc) = \vchannel(x)*c $\\ \hline
  \end{tabulary}
  \caption{}
  \label{tab:math:artist:nu}
\end{table}

The Stevens measurement types\cite{stevensTheoryScalesMeasurement1946}, listed in \autoref{tab:math:artist:nu}, are specified in terms of groups, which are monoids with invertible operations\cite{remlingAlgebraMath5353}. Despite critiques of the scales\cite{johnsonPseudoMathematicsMentalSocial1936,thomasMathematizationNotMeasurement2014}, we believe it is critical for the model to include the measurement scales since they are commonly used in visualization to classify components \cite{munznerVisualizationAnalysisDesign2014,wilkinsonGrammarGraphics2005}. By specifying the equivariance constraints on \vchannel\, we can guarantee that the stage of the artist that transforms data components into visual representations is equivariant. These constraints guide the implementation of reusable component transformers \vchannel\ that are composed when generating the graphic.

\subsubsection{Visualization Assembly}
\label{sec:math:artist:q}
\begin{figure}[htb]
  \centering
  \includegraphics[width=\textwidth]{figures/math/path_of_q.png}
  \caption{The transform functions $\vchannel_i$ convert data $\dsection_i\in \dtotal$ to visual characteristics $\vsection_i \in \vtotal$, then \vmark\ assembles $\vsection_i$ into a graphic $\gsection \in \gtotal$.} 
  \label{fig:math:artist:q}
\end{figure}
The transformation from data into graphic is analagous to a map-reduce operation; as illustrated in \autoref{fig:math:artist}, data components $\dtotal_{i}$ are mapped into visual components $\vtotal_{i}$ that are reduced into a graphic in \gtotal. The space of all graphics that \vmark\ can generate is the subset of graphics reachable via applying the reduction function $\vmark(\Gamma(\vtotal)) \in \Gamma(\gtotal)$ to the visual section $\vsection \in \Gamma(\vtotal)$. The full space of graphics is not necessarily equivariant; therefore we formalize the constraints on \vmark\ such that it produces structure preserving graphics.

@@

For example, given fiber $\vfiber=(xpos,\, ypos,\, color,\, thickness)$, then sections $\vsection=(0,0,0,1)$ and $\vmark(\vsection) = \gsection$ generates a piece of the thin hollow circle. The action $m=(e, e, e, x+2)$, where e is identity, translates \vsection\ to  $\vsection^{\prime}=(e,e,e,3)$ and the corresponding action on \gsection\ causes $\vmark(\vsection^{\prime})$ to be the thicker circle in \autoref{fig:math:artist:graphic}.

We formally describe a glyph as \vmark\ applied to the regions \dbasepoint\ that map back to a set of path connected components $\dbasepath \subset \dbase$ as input 
\begin{equation}
\dbasepath = \{\dbasepathpoint \in \dbase \text{ s. t. } \exists \gamma \text{ s.t. } \gamma(0)=\dbasepoint \text{ and }\gamma(1)=\dbasepathpoint\}
\end{equation}
where the path\cite{ConnectedSpace2020}  $\gamma$ from \dbasepoint\ to \dbasepathpoint\ is a continuous function from the interval [0,1]. We define the glyph as the graphic generated by $\vmark(\gbase_{\dbasepathpoint})$
\begin{equation}
  \begin{tikzcd}
      \gtotal \arrow[r, shift left] & \gbase_\dbasepathpoint \arrow[rr, "\vindex(\gbasepoint)", shift left] \arrow[l, "\gsection(\gbase_\dbasepathpoint)"] &  & \dbasepath_{\dbasepoint} \arrow[ll, "\vindex^{-1}(\dbasepath)"]
      \end{tikzcd}
  \label{eq:mark}
\end{equation}
such that for every glyph there is at least one corresponding region on \dbase, in keeping with the definition of glyph as any visually differentiable element put forth by Ziemkiewicz and Kosara\cite{ziemkiewiczEmbeddingInformationVisualization2009}. The primitive point, line, and area marks\cite{bertinSemiologyGraphicsDiagrams2011a,carpendaleVisualRepresentationSemiology} are specially cased glyphs.

\subsubsection{Assembly \vmark}
Given the continuities described in \ref{fig:math:graphic:retraction:map}, 
we illustrate a minimal Q that will generate the most minimal visualizations associated with those continuities: non-overlapping scatter points, a non-infinitely thin line, and an image. 
\begin{figure}[H]
    \includegraphics[width=1\textwidth]{figures/math/scatter.png}
    \caption{The data is discrete points (precipitation, time). Via \vchannel\ these are converted to (xpos, ypos) and pulled over discrete \gbase. These values are then used to parameterize \gsection\ which returns a color based on the parameters (xpos,ypos) and position $\alpha, \beta$ on $\gbase_{\dbasepoint}$ that \gsection\ is evaluated on. 
    }
    \label{fig:math:artist:q:scatter}
\end{figure}
The scatter plot in \autoref{fig:math:artist:q:scatter} has a constant size and color $\gsection_{RGB} = (0,0,0)$ that are defined as part of the point assembly function.

@@

\begin{figure}[H]
    \includegraphics[width=\textwidth]{figures/math/line.png}
    \caption{The line fiber $(time,\, temp)$ is thickened with the derivative $(time^{\prime},\, precipitation^{\prime}$ because that information will be necessary to figure out the tangent to the point to draw a line. This is because the line needs to be pushed perpendicular to the tangent of (xpos, ypos). The data is converted to visual characteristics (xpos, ypos). The $\alpha$ coordinates on \gbase\ specifies the position of the line, the $\beta$ coordinate specifies thickness.}
    \label{fig:math:artist:q:line}
\end{figure}
In contrast, the line plot in \autoref{fig:math:artist:q:line} has a \vindex\ function that is not only parameterized on \dbasepoint\ but also on the $\alpha$ distance along \dbasepoint\ and corresponding region in \gbase.

@@

\subsubsection{ \vmarkd}
\label{sec:math:artist:qhat}

\begin{figure}[htb]
    \centering
      \includegraphics[width=1\textwidth]{figures/math/q_hat.png}
      \caption{Because the pullback of the visual bundle \vtotalpull\ is the replication of a \vsection\ over all points \gbasepoint\ that map back to a single \dbasepoint, we can constract a \vmarkd\ on \vsection\ over \dbasepoint\ that will fabricate the \vmark\ for the equivalent region of \gbasepoint\ associated to that \dbasepoint}
      \label{fig:math:artist:qhat}
  \end{figure}

>>>>

\subsection{Artist}
\label{sec:artist}
We propose that visualization is structure preserving maps from data \dtotal\ to graphic \gtotal; having described \dtotal in \autoref{sec:math:data} and \gtotal in \autoref{sec:math:graphic}, we now define the visual transformations from \dtotal\ to \gtotal\ that formalize the components that visualization libraries implement. The topological artist \vartist\ is a map from the sheaf on a data bundle \dtotal\ which is $\mathcal{O}(\dtotal)$ to the sheaf on the graphic bundle \gtotal, $\mathcal{O}(\gtotal)$. 
\begin{equation}
    \vartist: \mathcal{O}(\dtotal) \rightarrow \mathcal{O}(\gtotal)
    \label{eq:math:artist:artist}
\end{equation}
The artist preserves \textit{continuity} through the \vindex\ map discussed in \autoref{sec:math:graphic:base}.  We propose that the artist $\mathscr{\vartist}$ is an \textit{equivariant} map of monoid action  $m \in \monoid$
\begin{equation}
\vartist(m\cdot \delement) = \varphi(m)\cdot \vartist(\delement) 
\end{equation}
between data element $\delement \in \mathscr{\dtotal}$ and graphic element $\vartist(\delement) \in \mathscr{\gtotal}$. To be equivariant with respect to monoids action, we conjecture that an  artist carries a monoid homomorphism $\varphi$
\begin{equation}
    \varphi: \monoid \rightarrow \monoid^{\prime}
\end{equation}
such that an action in data space $m \in \monoid$ is equivalent to an action in graphic space $\varphi(\monoid) \in \monoid^{\prime}$.

The artist \vartist\ has two stages: the encoders $\vchannel:\dtotal^{\prime} \rightarrow \vtotal$ convert the data components to visual components, and the assembly function $\vmark: \vtotalpull \rightarrow \gtotal$ composites the fiber components of \vtotalpull\ into a graphic in \gtotal.
\begin{equation}
    \label{eq:math:artist:diagram}
    \begin{tikzcd}
        \dtotal \arrow[r, "\vchannel"] \arrow[rd, "\pi"'] & \vtotal \arrow[d, "\pi"] & \vindex^*\vtotal \arrow[r, "\vmark"] \arrow[d, "\vindex^*\pi"'] \arrow[l, "\vindex^*"'] & \gtotal \arrow[ld, "\pi"] \\
                                              & \dbase                  & \gbase \arrow[l, "\vindex"']                                              &                    
        \end{tikzcd}
\end{equation}
\vtotalpull\ is the visual bundle \vtotal\ pulled back over \gbase\ via the equivariant continuity map $\vindex:\gbase \rightarrow \dbase$ introduced in \autoref{sec:math:graphic:base}.
The functional decomposition of the visualization artist in \autoref{eq:math:artist:diagram} facilitates building reusable components at each stage of the transformation because the equivariance constraints are defined on \vchannel, \vmark, and \vindex. We name this map the artist as that is the analogous part of the  Matplotlib\cite{hunterArchitectureOpenSource} architecture that builds visual elements.

\subsubsection {Visual Fiber Bundle \vtotal}
\label{sec:math:visual}
We introduce a visual bundle \vtotal\ to store the mappings of the data components into components of the graphic. These graphic components are implicit visualization library APIS; by making them explicit as components of the fiber we can define expectations of how these parameters behave. As with the data and graphic bundles, the visual bundle $(\vtotal,\,\dbase,\,\pi ,\,\vfiber)$ is defined by the projection map $\pi$
\begin{equation}
    \begin{tikzcd}[ampersand replacement=\&]
        \vfiber \arrow[r, hook] \& \vtotal \arrow[d, "\pi"'] \\
                          \& \dbase \arrow[u, "\vsection"', bend right]
    \end{tikzcd}
\end{equation}
where \vsection\ is the visual variable encoding, as described by Bertin \cite{bertinSemiologyGraphicsDiagrams2011a}, of the data section \dsection. The visual bundle \vtotal\ is the full design space\cite{cardStructureInformationVisualization1997} of possible parameters of a visualization type, such as a scatter plot or line plot. For example, one section \vsection\ of \vtotal\ is a tuple of visual values that specifies the visual characteristics of a part of a graphic.

\begin{table}[H]
    \centering
    \renewcommand{\arraystretch}{2}
    \begin{tabulary}{\textwidth}{|l|L|l|}\hline
     $\bm{\vchannel_{i}}$                      & $\bm{\vsection_{i}}$                                                            & $\bm{codomain(\vchannel_{i}) \subset \vfiber_{i}}$  \\ \hline                                              
    position                    & x, y, z, theta, r                                                          & $\mathbb{R}$   \\ \hline
    size                        & linewidth, markersize                                            & $\mathbb{R}^{+}$   \\ \hline
    shape                       & markerstyle                                                      & $\{f_{0}, \ldots, f_{n}\}$ \\ \hline
    color                       & color, facecolor, markerfacecolor, edgecolor  & $\mathbb{R}^{4}$ \\ \hline
    \multirow{2}{*}{texture}    & hatch                                                            & $\mathbb{N}^{10}$\\\cline{2-3}
                                & linestyle                                                        & $(\mathbb{R}, \mathbb{R^+}^{n, n\%2=0})$ \\ \hline              
    \end{tabulary}
    \caption{Some possible components of the fiber \vfiber\ for a visualization function implemented in Matplotlib}
    \label{tab:math:artist:mpl:fiber}
\end{table}
In \autoref{tab:math:artist:mpl:fiber}, the fiber components are specified by the visual parameter they are encoding. Multiple parameters can be encoded with the same transformation from data space to graphic space, for example x and y are both positions on a screen. Given a fiber of \(\{x, y, color\}\) one possible section could be  \(\{.5, .5, (255, 20,147)\}\). The \(codomain(\vchannel_i)\) in \autoref{tab:math:artist:mpl:fiber} specifies the libraries internal representation of visual variables and can be used to determine which monoids can act on $\vfiber_i$.

\subsubsection{Visual Encoders \vchannel}
\label{sec:math:artist:nu}
We propose that the map from data components to graphic components $\vchannel: \dsection \mapsto \vsection$ is a monoid \textit{equivariant} map. By specifying this constraints, we can guarantee that the stage of the artist that transforms data components into graphic representations is equivariant. These constraints then guide the implementation of reusable component transformers \vchannel\ that are composed when generating the graphic. We define the visual transformers \vchannel\ 
\begin{equation}
  \label{eq:math:artist:nu}
  \{\vchannel_{0}, \ldots, \vchannel_{n}\}: \{\dsection_{0}, \ldots, \dsection_{n}\} \mapsto \{\vsection_{0}, \ldots, \vsection_{n}\}
\end{equation}
as the set of equivariant maps $\vchannel_i: \dsection_i \mapsto \vsection_i$. Given $\monoid_i$ is the monoid action on $\dtotal_i$ and that there is a monoid ${\monoid_i}^{\prime}$ on $\vtotal_i$, then there is a monoid homomorphism from $\varphi:\monoid_i \rightarrow {\monoid_i}^{\prime}$ that \vchannel\ must preserve. As mentioned in \autoref{sec:math:data:monoid}, monoid actions define the structure on the fiber components and are therefore the basis for equivariance. Therefore, a validly constructed \vchannel\ is one where the diagram of the monoid transform $m$ commutes
\begin{equation}
  \label{eq:math:artist:nu_commute}
\begin{tikzcd}
  \dtotal_i \arrow[r] \arrow[r, "\vchannel_i"] \arrow[d, "m_{\delement}"'] & \vtotal_i \arrow[d, "m_{\velement}"] \\
  \dtotal_i \arrow[r, "\vchannel_i"]                           & \vtotal_i               
\end{tikzcd}
\end{equation}
such that applying equivariant monoid actions to $\dtotal_{i}$ and $\vtotal_{i}$ preserves the map $\vchannel_{i}: \dtotal_{i}\rightarrow \vtotal_{i}$. In general, the data fiber $\dfiber_{i}$ cannot be assumed to be of the same type as the visual fiber $\vfiber_{i}$ and the actions of \monoid\ on $\dfiber_{i}$ cannot be assumed to be the same as the actions of $\monoid^{\prime}$ on \vfiber; therefore an equivariant $\vchannel_i$ must satisfy the constraint  
\begin{equation}
\vchannel_i(m_{\delement}(\dtotal_i)) = \varphi(m_{\delement})(\vchannel_i(\dtotal_i))
\label{eq:math:artist:nu:equivariance}
\end{equation} 
such that $\varphi$ maps a monoid action on data to a monoid action on visual elements. However, 
without a loss of generality we can assume that an action of \monoid\ acts on $\dfiber_{i}$ and on $\vfiber_{i}$ compatibly such that $\varphi$ is the identity function. We can make this assumption because we can construct a monoid action of \monoid\ on $\vfiber_{i}$ that is compatible with a monoid action of \monoid\ on $\dfiber_{i}$. We can then compose the monoid actions on the visual fiber $\monoid^{\prime} \times \vfiber_{i} \rightarrow \vfiber_{i}$ with the homomorphism $\varphi$ that takes \monoid\ to $\monoid^{\prime}$. This allows us to define a monoid action on \vfiber\ of \monoid\ that is $(m, \velement) \rightarrow \varphi(m)\bullet\velement$, which lets us incorporate $\varphi$ into the action $\bullet$ such that $\varphi$ does not need to be explicitly defined in the constraints.

\begin{figure}[htb]
  \centering
  \begin{subfigure}{.49\textwidth}
    \includegraphics[width=1\textwidth]{figures/math/partial_fixed.png}
    \caption{}
    \label{fig:math:artist:nu:valid}
  \end{subfigure}
  \begin{subfigure}{.49\textwidth}
    \includegraphics[width=1\textwidth]{figures/math/partial_invalid.png}
    \caption{}
    \label{fig:math:artist:nu:invalid}
  \end{subfigure}
  \caption{ The \vchannel\ mapping in \autoref{fig:math:artist:nu:valid} represented by the colored arrows is monotonic, and therefore monoid equivariant,  since $\nu(storm)=\nu(rain)$ satisfies the condition $\nu(storm) \geq \nu(storm)$. In contrast, the map from data component to visual component in \autoref{fig:math:artist:nu:invalid} is not monotonic, and therefore not monoid equivariant, because $rain \geq storm$ is mapped to elements with the reverse ordering $\nu(storm) \geq \nu(storm)$.}
  \label{fig:math:artist:nu}
\end{figure}
The mapping from weather state to umbrellla in \autoref{fig:math:artist:nu:valid} is monotonic, and therefore we conjecture equivariant, because $\vchannel(rain) = \vchannel(storm)$ satisfies the monotonic condition of $rain \geq storm$. \autoref{fig:math:artist:nu} is an example of how the model supports partially ordered data components, which was a motivation for defining equivariance as monoid homomorphisms. In contrast, the translation from weather state data to visual representation as umbrella emoji in \autoref{fig:math:artist:nu:invalid} is an invalid visual encoding map \vchannel\ because it is not monotonic and therefore not equivariant This is because the monotonic condition \(rain \geq storm \implies \vchannel(rain) \geq  \vchannel(storm)\) is not met since \(\vchannel(rain) \leq\vchannel(storm)\). To satisfy the monotonic condition for $rain \geq storm$, either red arrow in \autoref{fig:math:artist:nu:invalid} would have to go in a different direction.

\begin{table}[H]
\centering
  \renewcommand{\arraystretch}{2}
  \begin{tabulary}{\columnwidth}{|llL|}\hline
      \textbf{scale} & \textbf{group} & \textbf{constraint}\\ \hline
      nominal & permutation &  $\text{if } \delement_1 \neq \delement_2 \text{ then } \vchannel (\delement_1) \neq\vchannel(\delement_2)$\\
      ordinal &  monotonic & $\text{if } \delement_1 \leq \delement_2 \text{ then } \vchannel (\delement_1) \leq \vchannel(\delement_2)$\\
      interval &  translation &  $\vchannel (x + c) = \vchannel(x) + c$ \\
      ratio &  scaling &  $\vchannel(xc) = \vchannel(x)*c $\\ \hline
  \end{tabulary}
  \caption{Equivariance constraints for the Stevens' measurement scales\cite{MeasurementScalesStatistics}}
  \label{tab:math:artist:nu}
\end{table}

The Stevens measurement types\cite{stevensTheoryScalesMeasurement1946}, listed in \autoref{tab:math:artist:nu}, are specified in terms of groups, which are monoids with invertible operations\cite{remlingAlgebraMath5353}. We generalize to monoids to account for limitations in the types of data that can be described with the Stevens' scales \cite{johnsonPseudoMathematicsMentalSocial1936,thomasMathematizationNotMeasurement2014}

\subsubsection{Visualization Assembly}
Having described the maps to components in \autoref{sec:math:artist:nu}, we now specify the assembly function \vmarkd\ that composites components in \vtotal\ into a graphic in \gtotal. Since the component transformes \vchannel\ are equivariant, the equivariance constraints carry through to \vmarkd. We specify these constraints to guide the implementation of library components responsible for generating graphics.

\label{sec:math:artist:q}
\begin{figure}[htb]
  \centering
  \includegraphics[width=\textwidth]{figures/math/path_of_q.png}
  \caption{The transform functions $\vchannel_i$ convert data $\dsection_i\in \dtotal$ to visual characteristics $\vsection_i \in \vtotal$. These visual components $\vsection_i$ are then assembled by \vmark\ into a graphic $\gsection \in \gtotal$.} 
  \label{fig:math:artist:q}
\end{figure}
The transformation from data into graphic is analagous to a map-reduce operation; as illustrated in \autoref{fig:math:artist:q}, data components $\dtotal_{i}$ are mapped into visual components $\vtotal_{i}$ that are reduced into a graphic in \gtotal. The space of all graphics that \vmark\ can generate is the subset of graphics reachable via applying the reduction function $\vmark(\Gamma(\vtotal)) \in \Gamma(\gtotal)$ to the visual section $\vsection \in \Gamma(\vtotal)$. The full space of graphics is not necessarily equivariant; therefore we formalize the constraints on \vmark\ such that it produces structure preserving graphics.

@@

For example, given fiber $\vfiber=(xpos,\, ypos,\, color,\, thickness)$, then sections $\vsection=(0,0,0,1)$ and $\vmark(\vsection) = \gsection$ generates a piece of the thin circle. The action $m=(e, e, e, x+2)$, where e is identity, translates \vsection\ to  $\vsection^{\prime}=(e,e,e,3)$ and the corresponding action on \gsection\ causes $\vmark(\vsection^{\prime})$ to be the thicker circle in \autoref{fig:math:artist:graphic}.

We formally describe a glyph as \vmark\ applied to the regions \dbasepoint\ that map back to a set of path connected components $\dbasepath \subset \dbase$ as input 
\begin{equation}
\dbasepath = \{\dbasepathpoint \in \dbase \text{ exists } \gamma \text{ s.t. } \gamma(0)=\dbasepoint \text{ and }\gamma(1)=\dbasepathpoint\}
\end{equation}
where the path\cite{ConnectedSpace2020}  $\gamma$ from \dbasepoint\ to \dbasepathpoint\ is a continuous function from the interval [0,1]. We define the glyph as the graphic generated by $\vmark(\gbase_{\dbasepathpoint})$
\begin{equation}
  \begin{tikzcd}
      \gtotal \arrow[r, shift left] & \gbase_\dbasepathpoint \arrow[rr, "\vindex(\gbasepoint)", shift left] \arrow[l, "\gsection(\gbase_\dbasepathpoint)"] &  & \dbasepath_{\dbasepoint} \arrow[ll, "\vindex^{-1}(\dbasepath)"]
      \end{tikzcd}
  \label{eq:mark}
\end{equation}
such that for every glyph there is at least one corresponding region on \dbase, in keeping with the definition of glyph as any visually distinguishable element put forth by Ziemkiewicz and Kosara\cite{ziemkiewiczEmbeddingInformationVisualization2009}. The primitive point, line, and area marks\cite{bertinSemiologyGraphicsDiagrams2011a,carpendaleVisualRepresentationSemiology} are specially cased glyphs.

\subsubsection{Assembly \vmark}
Given the continuities described in \ref{fig:math:graphic:retraction:map}, 
we illustrate a minimal Q that will generate the most minimal visualizations associated with those continuities: non-overlapping scatter points, a non-infinitely thin line, and an image. 
\begin{figure}[H]
    \includegraphics[width=1\textwidth]{figures/math/scatter.png}
    \caption{The data is discrete points (time, precipitation). Via \vchannel\ these are converted to (xpos, ypos) and pulled over discrete \gbase\ via $\vindex^*$. The pulled back visual secion $\vchannel$ is composited with the assembly function $\vmarkd \circ  \vchannel=\gsection$ to produce the instructions to make the graphic \gsection. The graphic section fills in the pixels in the screen via lookup on \gbase.}
    \label{fig:math:artist:q:scatter}
\end{figure}
The scatter plot in \autoref{fig:math:artist:q:scatter} has a constant size and color $\gsection_{RGB} = (0,0,0)$ that are defined as part of the point assembly function.

@@

\begin{figure}[H]
    \includegraphics[width=\textwidth]{figures/math/line.png}
    \caption{The line fiber $(time,\, precipitation)$ is thickened with the derivative $(time^{\prime},\, precipitation^{\prime}$ because that information will be necessary to figure out the tangent to the point to draw a line. This is because the line needs to be pushed perpendicular to the tangent of (xpos, ypos). The data is converted to visual characteristics (xpos, ypos). The $\alpha$ coordinates on \gbase\ specifies the position of the line, the $\beta$ coordinate specifies thickness.}
    \label{fig:math:artist:q:line}
\end{figure}
In contrast, the line plot in \autoref{fig:math:artist:q:line} has a \vindex\ function that is not only parameterized on \dbasepoint\ but also on the $\alpha$ distance along the interval \dbasepoint\ and corresponding region in \gbase.

@@

\subsubsection{Assembly Template\vmarkd}
\label{sec:math:artist:qhat}
The graphic base space \gbase\ is not accessible in many architectures, including Matplotlib; instead we can construct a factory function \vmarkd\ over \dbase\ that can build a \vmark. As shown in \autoref{eq:math:artist:diagram}, \vmark\ is a bundle map $\vmark: \vtotalpull\rightarrow \gtotal$ where \vtotalpull\ and \gtotal\ are both bundles over \gbase.
%[column sep=huge]
\begin{equation}
    \begin{tikzcd}
        \textcolor{gray!50}{\dtotal} \arrow[r, "\vchannel", color=gray!50] \arrow[rd, "\pi"', color=gray!50] & \vtotal \arrow[d, "\pi"']                  & \vindex^*\vtotal \arrow[r, "\vmark", color=gray!50] \arrow[d, "\vindex^*\pi"'] \arrow[l,  "\vindex^*"'] & \textcolor{gray!50}{\gtotal} \arrow[ld, "\pi", color=gray!50] \\
                                                          & \dbase \arrow[u, "\vsection"', bend right] & \gbase \arrow[l, "\vindex"'] \arrow[u, "\vindex^*\vsection"', bend right]               &                          
        \end{tikzcd}
        \label{eq:math:artist:qhat}
\end{equation}
